---
title: "Data in PDF files"
author: Pablo Barbera
output: html_document
---

#### Extracting text from PDF files

The other common scenario consists on extracting text that is embedded in PDF files. As noted earlier, how easy it is to convert the PDF file into machine-readable text will depend on whether the text is internally stored as such, and not as an image.

There are different methods to extract the text. Here I'll show the one that in my opinion is better -- `pdftotext`, an open-source tool that is part of the [Xpdf project](http://www.foolabs.com/xpdf/about.html). If you want to install it in your laptop, you can [download it here](http://www.foolabs.com/xpdf/download.html). It should work on all operating systems. See the `INSTALL` files for specific instructions.

We will run `pdftotext` not in the R console, but in your operating system's console. For example, if you're using a Mac, you would open the terminal and type the code. Since the way to do this varies across systems, we will instead run it from within R using the `system` function.

```{r}
system("pdftotext")
```

As you can see, we can ran pdftotext with different configurations. Which one is best will depend on your application. As a first example, let's look at a press release from the European Court of Human Rights about the outcome a case.

```{r}
system("pdftotext press-release.pdf")
```

The output file (in plain text) will have the same name unless we change it.

```{r}
system("pdftotext press-release.pdf press-release-output.txt")
```

If you look at the text of the file, you can see some of what we discussed earlier - any text that is internally represented as an image cannot be parsed.

We can also choose the specific pages to parse:

```{r}
system("pdftotext -f 1 -l 2 press-release.pdf")
```

Let's now work on a more advanced example. The document `arrests.pdf` contains a [list of Argentinian citizens arrested during the Pinochet regime](http://ftp2.errepar.com/bo/2013/04/SUPLE_17-04-2013.pdf). We'll try to parse the list here into a data frame format.

```{r}
system("pdftotext arrests.pdf")
```

Note that by default `pdftotext` will try to ignore the column layout, but if we wanted we would keep it:

```{r}
system("pdftotext -layout arrests.pdf arrests-layout.txt")
```

We can now use regular expressions to identify the blocks of text with the names of the arrests (because they are always in the first article of each decree), as well as the dates (because they all have "Bs. As." right before):

```{r}
txt <- readLines("arrests.txt", encoding = "latin1")
txt[200:250]
# names of those arrested
ar.init <- grep("Arréstese.*", txt)
ar.ends <- grep("Art. 2.*", txt)
txt[ ar.init[1] : ar.ends[1] ]
# dates of arrests
dates <- grep("^Bs\\. As\\.", txt)
txt[ dates[1] ]
```

Let's try to scrape the data for the first set of arrests and later on we'll generalize:

```{r}
init <- ar.init[1]
end <- ar.ends[ar.ends > init][1] # the first end line after the line we just chose
# this is what we will try to extract:
(data <- txt[init:end])
# now let's convert everything into a single string
data <- paste(data, collapse=" ")
# note that everything before "Ejecutivo Nacional" and starting with "Art. 2" is useless
data <- gsub(".*Nacional a:? (.*)\\. Art.*", data, repl="\\1")
# and let's split it back into substrings divided by ";"
data <- strsplit(data, ";")[[1]]
# we're almost there! Note that the name is everything *before* the parenthesis
names <- gsub(" ?(.*) \\(.*", data, repl="\\1")
# and the DNI (ID number) is everything *after*
dni <- gsub(".*\\((.*)\\)", data, repl="\\1")

# now, let's go back to the date
date <- txt[tail(dates[dates < init], n=1)] # first date after init
date <- gsub("Bs\\. As\\., ", "", date) # remove Bs. As.

# put everything into a data frame
df <- data.frame(date, names, dni, stringsAsFactors=F)

```

That seemed to work! Let's now replicate it for the entire dataset, inside a loop:

```{r}
arrests <- c()

for (init in ar.init){
	
	# extracting text
	end <- ar.ends[ar.ends > init][1]
	data <- txt[init:end]
	# cleaning text
	data <- paste(data, collapse=" ")
	data <- gsub(".*Nacional a:? (.*)\\. Art.*", data, repl="\\1")
	data <- strsplit(data, ";")[[1]]
	# extracting names and DNI
	names <- gsub("(.*) \\(.*", data, repl="\\1")
	dni <- gsub(".*\\((.*)\\)", data, repl="\\1")
	# extracting dates
	date <- txt[tail(dates[dates < init], n=1)] # first date after init
	date <- gsub("Bs\\. As\\., ", "", date) # remove Bs. As.
	# everything into a data frame
	df <- data.frame(date, names, dni, stringsAsFactors=F)
	arrests <- rbind(arrests, df)
}
```

If you look at the data frame, you'll see it's not completely perfect, but the rest we could edit them by hand, or tweak the function until it works.

