---
title: "Unsupervised text scaling"
author: Ken Benoit & Pablo Barbera
date: February 26, 2019
output: html_document
---

## Wordfish

The following code replicates the Irish 2010 budget debate scaling from Lowe and Benoit (2013)

```{r}
# loading data
library(quanteda)
data(data_corpus_irishbudget2010)
# making prettier document names
docnames(data_corpus_irishbudget2010) <-paste(docvars(data_corpus_irishbudget2010, "name"),
                                              docvars(data_corpus_irishbudget2010, "party"))
# creating DFM
ieDfm <- dfm(data_corpus_irishbudget2010)

# fitting wordfish
wf <- textmodel_wordfish(ieDfm, dir=c(2,1))

# looking at results
sort(predict(wf))

# extracting rest of parameters
str(coef(wf))

# let's look at the word-level parameters
sw <- data.frame( word=wf$features, beta=wf$beta, psi=wf$psi)
sw <- sw[order(sw$beta),] # discrimination
head(sw, n=20)
tail(sw, n=20)

sw <- sw[order(sw$psi),] # frequency
head(sw, n=20)
tail(sw, n=20)

# plotting both into an "Eiffel tower" plot
plot(wf$beta, wf$psi, type = "n", 
     xlab = "Word discrimination parameters", 
     ylab = "Word frequency parameters")
text(wf$beta, wf$psi, wf$features, col = "grey80", cex = .5)

```

### Wordfish applied to Twitter data

To explore an unsupervised approach to ideological scaling, let's continue with last week's example using tweets by Members of Congress. Can we recover a latent ideological dimension based on the text of their tweets?

```{r}
cong <- read.csv("congress-tweets.csv", stringsAsFactors=F)
# create corpus object
ccorpus <- corpus(cong$text)
docnames(ccorpus) <- cong$screen_name
# create DFM
cdfm <- dfm(ccorpus, remove_punct=TRUE, 
            remove=c(stopwords("english"), "t.co", "https", "rt", "amp", "http", "t.c", "can"))

# note heavy feature selection!
cdfm <- dfm_trim(cdfm, min_docfreq = 25)

# running wordfish
wf <- textmodel_wordfish(cdfm, dir=c(34, 42))

# how did we chose the two MCs that identify the model?
cong$screen_name[34]
cong$screen_name[42]

# looking at results
wf

# let's look at the word-level parameters
sw <- data.frame( word=wf$features, beta=wf$beta, psi=wf$psi)
sw <- sw[order(sw$beta),] # discrimination
head(sw, n=20)
tail(sw, n=20)

sw <- sw[order(sw$psi),] # frequency
head(sw, n=20)
tail(sw, n=20)

# plotting both into an "Eiffel tower" plot
plot(wf$beta, wf$psi, type = "n", 
     xlab = "Word discrimination parameters", 
     ylab = "Word frequency parameters")
text(wf$beta, wf$psi, wf$features, col = "grey80", cex = .5)

# and now we can compare the estimate positions with the ideal points...
plot(wf$theta, cong$idealPoint,
      xlab="Wordfish estimates", 
      ylab="Ideal points from roll-call votes",
      col=ifelse(cong$party=="R", "red", "blue"))
cor(wf$theta, cong$idealPoint)
cor(wf$theta[cong$party=="R"], cong$idealPoint[cong$party=="R"])
cor(wf$theta[cong$party=="D"], cong$idealPoint[cong$party=="D"])

```

## Wordfish vs correspondence analysis

Returning to the Lowe and Benoit (2013) example, we now fit it using correspondence analysis:

```{r}
# fitting wordfish
wf <- textmodel_wordfish(ieDfm, dir=c(2,1))
# fitting correspondence analysis
wca <- textmodel_ca(ieDfm, nd=2) # two dimensions

# looking at legislators' estimated positions
wca$rowcoord

# words' positions
head(wca$colcoord)

# comparing wordfish and CA results
plot(wf$theta, wca$rowcoord[,1], 
     xlab="Wordfish theta-hat", ylab="CA dim 1 coordinate",
     main="Wordfish v CA dim 1", pch=19, xlim = c(-1.5, 1.5), ylim = c(-1.5, 1.5))
abline(lm(wca$rowcoord[,1] ~ wf$theta), col="grey50", lty="dotted")
```

What happens if we add some irrelevant text?

```{r}

## demonstrate the effects of irrelevant text
newText <- texts(data_corpus_irishbudget2010)
# some really irrelevant text, from Justin Bieber's Wikipedia page
junkText <- "   
Justin Drew Bieber (born March 1, 1994)[4][5] is a Canadian singer and songwriter. Bieber's current manager, Scooter Braun first discovered him through his YouTube videos in 2007.[6] Braun was impressed with the videos, and contacted Bieber's mother Pattie about wanting to work with him. After convincing Bieber's mother, Braun arranged for Bieber to meet with Usher in Atlanta, Georgia. Bieber was signed to RBMG,[7] and then to an Island Records recording contract offered by record executive, L.A. Reid.[3][8]\n\n
Bieber released his debut EP, My World, in November 2009. It was certified platinum in the United States. He became the first artist to have seven songs from a debut record to chart on the Billboard Hot 100.[9] Bieber released his first full-length studio album, My World 2.0, in March 2010. It debuted at or near number one in several countries and was certified platinum in the United States. It was preceded by the single \"Baby\", which is also, as of July 2014, the YouTube video with the most dislikes, and the most views on a Vevo platform. He followed up the release of his debut album with his first headlining tour, the My World Tour, the remix albums My Worlds Acoustic and Never Say Never – The Remixes, and the 3D biopic-concert film Justin Bieber: Never Say Never. Bieber released his second studio album Under the Mistletoe in November 2011, when it debuted at number one on the Billboard 200. Bieber released his third studio album Believe in June 2012. It became his fourth chart topper in the United States, debuting at number-one on the Billboard 200.\n\n
Bieber has received numerous industry awards, including both Artist of the Year Awards, at the 2010 American Music Awards, and the 2012 American Music Awards, and was nominated for Best New Artist and Best Pop Vocal Album at the 53rd Grammy Awards. With a global fan base, termed as \"Beliebers\", and over 40 million followers on Twitter, he was named by Forbes magazine in 2012 as the third-most powerful celebrity in the world.[10] However Bieber fell to number 9 on the list in 2013.
"
newText[11] <- paste(newText[11], junkText)
ieDfm2 <- dfm(newText)
# fit wordfish again
wf2 <- textmodel_wordfish(ieDfm2, dir=c(2,1))
# plot it and compare Ryan
theta1 <- wf$theta; theta2 <- wf2$theta
names(theta1) <- names(theta2) <- wf$docs
sortOrder <- order(theta1)
theta1 <- theta1[sortOrder]
theta2 <- theta2[sortOrder]
par(mfrow=c(2,1))
dotchart(theta1, main="Without irrelevant text", col = c(rep("black", 11), "red", rep("black", 2)))
dotchart(theta2, main="With irrelevant text",    col = c(rep("black", 11), "red", rep("black", 2)))


```

